\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{tabularx}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{threeparttable}
\usepackage{array}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{siunitx}
  \sisetup{
    table-format=1.3,
    round-mode = places,
    round-precision = 3,
  }
\usepackage{kantlipsum}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Homework 2 - A Study on Privacy-centered Web Browsers\\
}

\author{\IEEEauthorblockN{Ishaan Sathaye}
\IEEEauthorblockA{\textit{CSC 325 Intro to Privacy} \\
\textit{California Polytechnic State University, San Luis Obispo}\\
San Luis Obispo, U.S. \\
isathaye@calpoly.edu}
}

\maketitle

\begin{abstract}
This paper is a summary of 2 privacy-centered web browsers, which are Tor and
Brave. In addition, this paper also includes a comparison of the privacy
preserving features of these browsers.
\end{abstract}

\section{Privacy-Centered Browsers}

\subsection{Tor}

Built on the Firefox platform, the Tor Browser is a privacy-focused web browser
designed to enhance user anonymity on the web. My motivation for choosing this
browser is that it provides a secure and private browsing experience by routing
all the internet traffic through its Tor network. This network is a series of
nodes that are run by volunteers, which makes it difficult to track the
activities of the users. These nodes encrypt and anonymize the traffic by
pushing user data through multiple relays, making it very challenging for anyone
to track a person's internet activity. This browser includes a variety of
features such as a built-in ad blocker, HTTPS Everywhere, NoScript
integration or script control, and regular updates to enhance the security of
the browser\cite{b1}. All in all, this browser ois prefect for people seeking a secure
online privacy, especially in places where anonymity is a concern.

\subsection{Brave}

Built on the Chromium platform, the Brave Browser is a privacy-focused web
browser that stands out ofr its robust ad and tracker blocking features. My
motivation for choosing this browser is its utilization of "shields" to reduce
users' exposure to online tracking, which enhances overall privacy. This browser
gives the users the power to control their privacy settings, which enables them
to block scripts, fingerprinting, and other tracking methods. With the automatic
enforcement of HTTPS and the option to use the Tor network, this browser
is versatile and can be used by anyone seeking a secure and private browsing
experience. Its unique Basic Attention Token (BAT) system also allows users to
opt in to view ads and earn tokens, which is a great incentive for people to use
this browser\cite{b2}. Respecting advertising and earning tokens for viewing ads is an
unique feature that makes this browser stand out from the rest.

\section{Privacy-Preserving Features}

\begin{table*}[h]
    \renewcommand{\arraystretch}{1.3}
    \caption{Comparison of Privacy-Preserving Features}
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
    \textbf{Feature} & \textbf{Tor Browser} & \textbf{Brave Browser} \\
    \hline
    Tor Network & \checkmark & Limited \\
    \hline
    HTTPS Everywhere & \checkmark & \checkmark \\
    \hline
    NoScript Integration & \checkmark & \checkmark \\
    \hline
    Regular Updates & \checkmark & \checkmark \\
    \hline
    Built-in Ad Blocker & \checkmark & \checkmark \\
    \hline
    Tor Circuit & \checkmark & \checkmark \\
    \hline
    Deep Anonymity & \checkmark & \\
    \hline
    Decentralized Relays & \checkmark & \\
    \hline
    Government Censorship Resistance & \checkmark & \\
    \hline
    Shields & & \checkmark \\
    \hline
    Automatic HTTPS Enforcement & & \checkmark \\
    \hline
    Brave Rewards & & \checkmark \\
    \hline
    Brave Search & & \checkmark \\
    \hline
    Brave Today & & \checkmark \\
    \hline
    Brave Firewall + VPN & & \checkmark \\
    \hline
    \end{tabular}
\end{table*}

\subsection{Tor's Privacy-Preserving Features}

Tor's privacy-preserving features include the following\cite{b3}:

\begin{itemize}
    \item \textbf{Tor Network:} The Tor network is a series of nodes that are
    run by volunteers, which makes it difficult to track the activities of the
    users. These nodes encrypt and anonymize the traffic by pushing user data
    through multiple relays, making it very challenging for anyone to track a
    person's internet activity. In fact, Tor is short for "The Onion Router",
    which refers to the layers of encryption that are used to anonymize the
    traffic.
    \item \textbf{HTTPS Everywhere:} This feature ensures that the user's
    connection to the website is secure by encrypting the traffic. This is
    especially useful when the user is browsing on a public network.
    \item \textbf{NoScript Integration:} This feature allows the user to control
    the scripts that are allowed to run on the browser. This is useful for
    blocking malicious scripts that can be used to track the user's activity.
    \item \textbf{Regular Updates:} This feature ensures that the browser is
    up-to-date with the latest security patches, which enhances the security of
    the browser.
    \item \textbf{Built-in Ad Blocker:} This feature blocks ads and trackers,
    which enhances the user's privacy. This is especially useful for blocking
    malicious ads that can be used to track the user's activity.
    \item \textbf{Tor Circuit:} This feature allows the user to view the
    relays that are used to anonymize the traffic. This is useful for
    understanding how the Tor network works.
\end{itemize}

\subsection{Brave's Privacy-Preserving Features}

Brave's privacy-preserving features include the following\cite{b4}:

\begin{itemize}
    \item \textbf{Shields:} This feature allows the user to control their
    privacy settings, which enables them to block scripts, fingerprinting, and
    other tracking methods. This is useful for blocking malicious scripts that
    can be used to track the user's activity.
    \item \textbf{Tor Network:} This feature allows the user to use the Tor
    network to enhance their privacy. This is useful for people seeking a secure
    and private browsing experience.
    \item \textbf{Automatic HTTPS Enforcement:} This feature makes sure that 
    the user's connection to the website is secure by encrypting the traffic.
    \item \textbf{Regular Updates:} Enhances the security of the browser
    by ensuring that the browser is up-to-date with the latest security patches.
    \item \textbf{Built-in Ad Blocker:} This feature blocks ads and trackers,
    which enhances the user's privacy. This is especially useful for blocking
    malicious ads that can be used to track the user's activity.
    \item \textbf{Brave Rewards:} This feature allows the user to opt in to
    view ads and earn tokens, which is a great incentive for people to use this
    browser. Respecting advertising and earning tokens for viewing ads is an
    unique feature that makes this browser stand out from the rest.
    \item \textbf{Brave Search:} This feature allows the user to use the Brave
    search engine, which is a privacy-preserving search engine that does not
    track the user's activity. This is useful for people seeking a secure and
    private browsing experience.
    \item \textbf{Brave Today:} This feature allows the user to view the latest
    news, which enables users to stay up-to-date with the latest news. This is
    great for people who want to stay informed about privacy-related news.
    \item \textbf{Brave Firewall + VPN:} This feature allows the user to use
    the Brave Firewall + VPN, which is a privacy-preserving VPN that does 
    \textbf{not} track the user's activity.
\end{itemize}

\section{User Experience}

\begin{table*}[h]
    \renewcommand{\arraystretch}{1.3}
    \caption{Comparison of User Experience}
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
    \textbf{Feature} & \textbf{Tor Browser} & \textbf{Brave Browser} \\
    \hline
    Installation & Straightforward & Straightforward \\
    \hline
    Setup & Straightforward & Straightforward \\
    \hline
    Interface & Firefox & Chrome \\
    \hline
    Responsiveness & Slow & Fast \\
    \hline
    Search Engine & DuckDuckGo & Default \\
    \hline
    Private Browsing Mode & \checkmark & \checkmark \\
    \hline
    UI & Unpolished & Polished \\
    \hline
    Accessability to Privacy Features & Easy & Medium \\
    \hline
    \end{tabular}
\end{table*}

\subsection{Tor's User Experience}

Tor's user experience is very similar to that of Firefox, which makes it easy
for people to use this browser. The installation of this browser is a very
straightforward process and upon opening the browser, the user is greeted with
a config wizard that guides users through the setup process. There is a familiar
Firefox interface such as the URL bar, bookmarks, and navigation tabs\cite{b5}. 
The default search engine is DuckDuckGo, which is a privacy-preserving search 
engine that does not track the user's activity. When the browser is launched, 
the user is connected to the Tor network and this process is very transparent to
the user. Browsing speed can sometimes be slow due to the multiple relays that 
are used to anonymize the traffic. There is still a private browsing mode that
allows the user to browse the web without saving any history.

From my own experience, the browser felt very slow and unresponsive at times,
and also the user interface was not very appealing. However, the privacy
features it offers are very useful. As I am a Safari user, this browser felt
very "heavy" as it comes with a lot of features that bloat software. However,
this browser is perfect for people seeking a secure and private browsing
experience.

\subsection{Brave's User Experience}

Installation of the Brave browser is done through its official website and it is
available for major operating systems such as Windows, macOS, Linux, Android,
and iOS. This setup process is also very straightforward and additionally users
can import their bookmarks, history, and passwords from other browsers such as
Chrome, Firefox, and Edge. The interface is very similar to that of Chrome,
which makes it easy for people to use this browser. There is an "omnibox" that
allows the user to search the web or enter a URL. Brave Shields is available
from the address bar and it allows users to access the privacy-preserving
features of the browser. Since this browser is built on the Chromium platform,
it is very fast and responsive \cite{b6}, and it also supports Chrome 
extensions. There is also a private browsing mode that allows the user to browse
the web without saving any history.

From my own experience, this browser was much faster and more responsive 
than Tor. Overall, the user interface was very appealing and it felt more 
polished than Tor. However, it can be confusing to find the privacy-preserving 
features of the browser as they are hidden in the address bar. This browser 
seemed more user friendly than Tor, as there was more emphasis on the user 
interface.


\section{Conclusion}

From this homework study, I learned about the privacy-preserving features of
Tor and Brave. I also learned about the user experience of these browsers and
how they compare to each other. Overall, I think that both of these browsers
are great for people seeking a secure and private browsing experience. However,
I think that Brave is more user friendly than Tor, as it has a more polished
user interface. Utility-wise I think that Tor is better than Brave, as it has
better privacy-preserving features such as the Tor network. In conclusion, I
think that both of these browsers are great for people seeking a secure and
private browsing experience, and it is up to the user to decide which browser
is better for them.

\begin{thebibliography}{00}
    \bibitem{b1} “SECURE CONNECTIONS | Tor Project | Tor Browser Manual.” n.d. Tb-Manual.torproject.org. \newline
    https://tb-manual.torproject.org/secure-connections/.
    \bibitem{b2} Keizer, Gregg. 2018. “The Brave Browser Basics - What It Does, How It Differs from Rivals.” Computerworld. July 24, 2018. \newline 
    https://www.computerworld.com/article/3292619/the-brave-browser-basics-what-it-does-how-it-differs-from-rivals.html.
    \bibitem{b3} Wherry, Jack . 2020. “What Is Tor (Browser) \& How Does It Work?” CyberNews. September 29, 2020. \newline
    https://cybernews.com/privacy/what-is-tor-and-how-does-it-work/.
    \bibitem{b4} “Features.” n.d. Brave Browser. https://brave.com/features/.
    \bibitem{b5} June 2017, Carrie Marshall 11. n.d. “Tor Browser Review.” TechRadar. https://www.techradar.com/reviews/tor-browser.
    \bibitem{b6} “A Beginner's Guide to the Privacy-Focused Next-Generation Brave Browser.” n.d. Cointelegraph. Accessed November 28, 2023. \newline
    https://cointelegraph.com/learn/a-beginners-guide-to-the-privacy-focused-next-generation-brave-browser.
\end{thebibliography}
    

\end{document}
